Visualizing Functional Data

### 2017-08-14 paper/sections/categorical.tex
close of rough categorical

%%find a citation that talks about categorical data in set theoretic terms

Categorical data can be likened to set measurment, where each category is a set
and researchers often want to know how many members each set has and what
intersections are common between sets. It is the latter that is analogous to a
conditional probability, the question of the likelihood of membership in one
set given membership in another. And sometimes there is also a quantative
component involved too. For example, in a study of how NEH funding has changed
over time, there are questions of the likelihoods of different disciplines
being funded (\{d_0,\dots,\d_i}\in \D), and if that is conditioned on institutione
  type \{i_0, \dots,\i_n}\in I}, and if either of those are then conditional on time \{t_0, \dots,t_n} \cite{wired} 
%% this can as easily be 311 data, but I don't have a citation for that
 The most common approach is usually to create a list of frequncies, sometimes of
 sets of one category, such as \D, or a crosstable wherein each cell is a an element
 in the product of two (or more sets), such as \D\cross \I creates a table
 wherein cell one is the frequency of elements that are in both $\d_0$ and
 $\i_0$ \cite{what is cross tabulation?}

@@

Another important factor in visualizing categorical data is retaining the
inherent heirachy in many datasets with categorical measurements
\cite{shneiderman_visualizing_2000}. The Parallel Sets tool
showin in figure~\ref{fig:upsetfig} is expicilitly designed to show frequencies
of set membership in heirarichal ordered categorical sets, so the \{d_0, \dots, d_n} \in D
  \cite{kosara_parallel_2006}. Building on the fexible linked axis version of
  the parallel cooridnates plot \cite{flexible linked axis}, Parallel Sets
  treats each categorical set independantly but indicates conditional
  dependency through cross tabulations and by grouping the data, indicated via
  color, based on an active dimension that then becomes the pivot point for the
  cross tabulations. In Parallel Sets, quantative data is converted to
  categorical via binning.

\begin{figure}

\end{figure}
Often though it's important to retain the internel structure of the
quantative data, such as when it has internal structucture such as time and
space. Zhang et al propose a method that retains data and allows for
visualizing more than just pairwise relationships\cite{zhang_visual_2015}. Thei

>>>>

Categorical data can be likened to set measurment, where each category is a set
and researchers often want to know how many members each set has and what
intersections are common between sets \cite{agresti_categorical_2011,schneider_set-theoretic_2012}
  It is the latter that is analogous to a
conditional probability, the question of the likelihood of membership in one
set given membership in another. And sometimes there is also a quantative
component involved too. The most common approach is usually to create a list of frequncies, sometimes of sets of one category, such as \D, or a crosstable wherein each cell is a an element
 in the product of two (or more sets), such as \D\cross \I creates a table
 wherein cell one is the frequency of elements that are in both $\d_0$ and
 $\i_0$.

@@

Another important factor in visualizing categorical data is retaining the
inherent heirachy in many datasets with categorical measurements
\cite{shneiderman_visualizing_2000}. The Parallel Sets tool
showin in figure~\ref{fig:upsetfig} is expicilitly designed to show frequencies
of set membership in heirarichal ordered categorical sets, so the \{d_0, \dots, d_n} \in D
  \cite{kosara_parallel_2006}. Building on the fexible linked axis version of
  the parallel cooridnates plot \cite{claessen_flexible_2011}, Parallel Sets
  treats each categorical set independantly but indicates conditional
  dependency through cross tabulations and by grouping the data, indicated via
  color, based on an active dimension that then becomes the pivot point for the
  cross tabulations. In Parallel Sets, quantative data is converted to
  categorical via binning.

\begin{figure}
  \includegraphics{corr.png}
  \label{caption} Zhang et al create an initial map (a) that they then apply an
  edge filter to (b). They then allow the user to filter out a variable and
  either highlight only the vertices of that variable (c) or the edges incident
  on that variable(d).
  \label{fig:corr}
\end{figure}

Often though it's important to retain the internel structure of the
quantative data, such as when it has internal structucture such as time and
space. Zhang et al propose a method that retains data and allows for
visualizing more than just pairwise relationships\cite{zhang_visual_2015}. They
introduce a concept of a correlation map, as shown in
figure~\ref{fig:corr}. The authors build out a graph where every variable is a
vertex and the edges are correlations between the connecting vertices. The
vertices are sized relative to the variance in the data. The edges are colored
red for negative correlations and green for positive correlations. To convert
the categorical data to numerical space, they developed an algorithm where when
mapped to numbers, the residual sum of squares of a logostic regression of the
categorical variables  would be minimized. They then filter edges with low
correlations and on variables of interest.
